A demonstration of the \LaTeXe\ class file for \itshape{SAGE Publications}

### 2015-12-02 body/contact.tex
finished contact generation now heuristics

\section{From a guide path to a discrete sequence of contact configurations (Stage 2)}
\label{sec:contact}
As an input of this stage, we consider a truly feasible guide path $\mathbf{q}^0(t) : [0,1] \longrightarrow SE(3)$ for the root of the robot $R$. We now consider the second problem of computing a discrete sequence of contact configurations $\mathbf{Q}^{\overline{0}}$ for the limbs of the robot along $\mathbf{q}^0(t)$. From true feasibility we know that such a trajectory exists.

@@

\subsection{Single contact generation}
We consider a limb $R^k$  and a root configuration $\mathbf{q}^0$ truly feasible regarding $R^k$. FIGURE
From true feasibility we know that there exists an infinite list of mappings $P = [\pi_0,..,\pi_i, ...]$ that verify Eq.~\eqref{eq:pi} for $\mathbf{q}^0$.
The issue is then to evaluate locally any element of $P$ at $\mathbf{q}^0$.
This can be achieved as follows:
\begin{enumerate}
\item Generate randomly a valid limb configuration;
\item Project the end-effector onto the closest surface with inverse kinematics;
\item If a feasible solution is found, stop. Otherwise repeat from step 1.
\end{enumerate}

We partially trade the completeness property of this approach for a more efficient solution \citep{Tonneau2014}:
\begin{enumerate}
\item Generate offline $N$ valid sample limb configurations;
\item using the end-effector position expressed in $R$ coordinates as an index, store each sample in an octree stata structure;
\item Figure TODO, 2 and 3: At runtime, when contact creation is required, retrieve from the octree the list of samples $S = [\mathbf{q}^k_0, ..., \mathbf{q}^k_i, ...]$ such that $dist(\mathbf{p}^k(\mathbf{q_k}^k),O) < \epsilon$, with $\epsilon $ small ;
\item According to a user-defined heuristic $h$, sort the elements of $S$;
\item FIGURE TODO, 4 and 5: Select the first element of $S$, and project the configuration to contact with inverse kinematics;
\item If a feasible solution is found, stop. Otherwise remove the element from $S$ and go from step 5, or fail if $S$ is empty.
\end{enumerate}

This approach has two main advantages:
First, this allows us to select a number of candidates at a small cost, (a single proximity request), and to sort them according to a heuristic;
Secondly, thanks to the use of proximity requests, the number of iterations necessary for the inverse kinematics to converge is reduced.
Finally, it is easy to see that as $N$ grows to infinity, the completeness of the method is restored.
$N$ is a parameter that allows us to specify the trade-off between completeness and efficiency.
The reader is invited to refer to our previous work for an extensive discussion on the optimal value of $N$ \citep{Tonneau2014}.

\subsubsection{Efficient contact generation }

Concretely, any, we can use a random sampling approach combined with an inverse kinematics projector.

given a root configuration $\mathbf{q}^0$ and a limb $R^k$, we locally identify a countable set of mappings $P = [\pi_0,..,\pi_i] $ that verify Eq.~\eqref{eq:pi} for $\mathbf{q}^0$. and select the most relevant according to a heuristic $h$.

%We compute the image of a mapping $f$ along $\mathbf{q}^0(t)$ into a discrete sequence of contacts $ \in C_{contact}$. We first describe the discretization of $\mathbf{q}^0(t)$ and the algorithm used to create or break contacts along it. Then we describe how contacts are explicitely generated along the discretized trajectory.

\subsection{Extension of the guide trajectory} 
The guide trajectory $\mathbf{q}^0(t)$ is first discretized into a sequence of $j$ key placements:  
\begin{equation*}
	\mathbf{Q}^0 = [\mathbf{q}^0_{0}; \mathbf{q}^0_{i}; ..., \mathbf{q}^0_{j-1}]
\end{equation*} 
where $\mathbf{q}^0_{0}$ and $\mathbf{q}^0_{j-1}$ respectively correspond to the start and goal configurations.  To ensure continuity in the contact transition phases, we rewrite $\pi$ under the following recursive form for any $0<i<j$:
\begin{equation*}
    \pi\colon\left\{
    \begin{aligned}		
        \mathbf{Q}^0 \in C_{reach} & \longrightarrow C_{contact} \\
        %\mathbf{q}^{0}_0 &  \longrightarrow  \mathbf{q}_{start} \\
        \mathbf{q}^{0}_i &  \longrightarrow  \mathbf{q}^{0}_i \oplus g(\mathbf{q}_{i - 1},\mathbf{q}^{0}_i) 
    \end{aligned}
    \right.
\end{equation*} 
We initialize the recurrence with $\pi(\mathbf{q}^0_{0}) = \mathbf{q}_0$ the initial configuration of the robot.
The function $g$ is defined independently by $g^k$ for each limb $R^k$. In defining $g^k$, two aspects must be considered. Is the limb $R^k$ in contact? And which criteria is it optimizing?

\medskip%\subsection{Maintaining and breaking contact in the sequence}
\noindent\textbf{Maintaining a contact:} If possible, a limb in contact at time $i-1$ remains in contact at $i$. The contact is broken if an inverse kinematics solver fails to find a collision free limb configuration which satisfies joint limits.
The solver is a direct implementation of the inverse kinematics solver with task priorities framework proposed in \cite{Baerlocher2004}, which
handles explicitely joint limits, and allows to integrate collision avoidance as a secondary criterion.

If the solver fails, the contact is broken and a collision free configuration is assigned to the limb.
%~ The configuration is selected to be the closest to the previous configuration (by minimizing $\| \mathbf{p}^k(\mathbf{q_i}) - \mathbf{p}^k(\mathbf{q_{i-1}}) \|$).

Once a first candidate configuration is taken for all limbs, the quasi-static balance is tested by whether the weight wrench is in the gravito-inertial cone (i.e. there exists valid contact forces that compensate for the weight of the robot), using the geometric approach described in \cite{qiu:dhm:2011}. If the balance is not obtained, new contacts are randomly generated using the following procedure.

%\begin{itemize} 
%\item If the robot is not statically balanced and $\mathbf{q}_{i-1}^k$ is not in contact, $g$ computes a configuration $\mathbf{q}_{i}^k$ so that a contact is taken with the environment (Fig.~\ref{fig:generate_contact}).
%\item Else, $g$ computes a configuration $\mathbf{q}_{i}^k$ minimizing $\| \mathbf{p}^k(\mathbf{q_i}) - \mathbf{p}^k(\mathbf{q_{i-1}}) \|$, subject to $R^k(\mathbf{q_i}) \cap O = \emptyset$ and $\mathbf{q_i}^k$ respects the  joint limits of $R^k$.
%
%If $\mathbf{q}_{i - 1}^k$ is in contact, this results in trying to maintain the contact for $\mathbf{q}_{i}^k$ (Figure~\ref{fig:break_contact}--1 and 2). The minimum is different than zero if $W^k(\mathbf{q}_i^0) \cap  \mathbf{p}^k(\mathbf{q_{i-1}}) = \emptyset$ (Figure~\ref{fig:break_contact}--3), or collision can not be  prevented (Figure~\ref{fig:break_contact}--4).
%\end{itemize}

\begin{figure}[t]
\centering
  \begin{overpic}[width=0.6\linewidth]{figures/break_contact}
		\put (3,15) {1)} 
		\put (27,15) {2)} 
		\put (50,15) {3)} 
		\put (75,15) {4)} 
		%~ \put (68,58) {3.a)} 
		%~ \put (5,27) {3.b)} 
		%~ \put (37,27) {4.a)} 
		%~ \put (68,27) {4.b)} 
	\end{overpic}
\caption{Contacts are maintained unless their position is too far, or the environment prevents it.}
		   \label{fig:break_contact}
\end{figure}

\begin{figure}[t]
\centering
  \begin{overpic}[width=0.6\linewidth]{figures/generate_contact}
		%~ \put (5,58) {1)} 
		%~ \put (37,58) {2)} 
		%~ \put (68,58) {3.a)} 
		%~ \put (5,27) {3.b)} 
		%~ \put (37,27) {4.a)} 
		%~ \put (68,27) {4.b)} 
	\end{overpic}
\caption{Contacts are generated when the configuration is not balanced.}
		   \label{fig:generate_contact}
\end{figure}

%\subsection{Generation of a contact configuration}  
\medskip\noindent\textbf{Creating a contact:} We consider a configuration where some limbs are in contact, some are free and quasi-static balance is not enforced. To enforce balance, we proceed in the following 
manner:
we randomly select a contact free limb; if there is no contact free limb, we select
the limb that made contact first.
Using the contact generator introduced in~\cite{Tonneau2014}, we project the configuration of this limb into a contact that enhances balance, if it exists (Figure~\ref{fig:generate_contact}); If balance is not achievable and a contact is possible, it is generated anyway;
If balance is not achieved, the next limb is selected and projected into a contact configuration, and so on.
This approach can lead to the repositioning of existing contacts, in which case intermediate states are inserted to reposition the contacts.
This current implementation does not guarantee that the planner will succeed in generating a balanced configuration, because
true feasibility is not fully guaranteed.
However in practice the planner is successful in the large majority of cases, as discussed in section~\ref{sec:stat1}.
%~ A new contact configuration is obtained by projecting a random configuration to the nearest contacts. The sampling is biased by selecting configurations so that the limbs previously in contact are already close to contact and at least one additional limb is close to the environment. The random configuration is then projected onto the contact by iteratively minimizing the distance between the end effectors and the environment, under the constraint of respecting the joint limits and avoiding the collision.

\subsection{Contribution to the global movement: the EFORT criteria}
\noindent\textbf{EFORT criterion:} If only relying on the random sampling to select new contacts, the planner produces inefficient postures. The resulting contact sequence is then poorly efficient and unnatural. Moreover, the limbs are not well configured and are not able to efficiently follow the general movement: contacts break frequently.

When creating additional contacts, we therefore propose to select particular configurations that allow to exert a force compatible with the direction of motion. This task efficiency is measured  with the Extended FORce Transmission ratio (EFORT)~\cite{Tonneau2014}.
%
The measure of EFORT is given by
\begin{equation}
\alpha_{EFORT}(\mathbf{q}^k, \mathbf{m}) = [\mathbf{m}^{T}(\mathbf{J}\mathbf{J}^T)\mathbf{m}]^{-\frac{1}{2}} ( \nu_0 \mathbf{n}^T \mathbf{m})
\end{equation}
where $\mathbf{J}$ is the Jacobian matrix of the limb $R^k$ in configuration $\mathbf{q}^k$; $\nu_0$ is the friction coefficient of the contact surface; $\mathbf{n}$ is the normal of the contact surface; and $\mathbf{m}$ is the direction opposite to the motion,
given by the 3D vector connecting $\mathbf{q}_{i}^0$ and $\mathbf{q}_{i+1}^0$.
%
The first part of the equation measures the ratio between the joint  torques and the resulting force applied along $\mathbf{m}$. The second part quantifies the odds of slipping while applying a force along $\mathbf{m}$.

\noindent\textbf{Optimization at creation:} In practice, a database of configurations is stored for each limb, which can be considered as manipulator arms. The database is implemented  as an octree data structure, indexed by the end-effector positions of the configurations (and additionally storing $\mathbf{J}$). 
%Fig.~\ref{fig:contact_gen}-2 presents a few configurations for the right arm of the robot.  
Upon request, the octree returns a set of configurations close to contact (Fig.~\ref{fig:contact_gen}-3). These candidates are sorted based on their task efficiency, given by $\alpha_{EFORT}$. The first candidate in this list satisfying the balance criterion and is collision free is selected and projected on the contact surface using our inverse kinematics solver.

%Therefore to generate a contact, we want to find a configuration which is collision free, allows to maintain balance, and maximizes EFORT. We choose this configuration among a set of possible candidates, rapidly computed thanks to a contact generator, also proposed in~\cite{Tonneau2014}. We recall the overall idea here. We consider each limb of the robot as a manipulator arm attached to the root. We store a database of configurations for each manipulator. The database is implemented  as an octree data structure, indexed by the end-effector positions of the configurations. The jacobian matrix $\mathbf{J}$ of each configuration is also stored. Figure~\ref{fig:contact_gen} -- 2 presents a few configurations for the right arm of the robot.

%To summarize: given a feasible trajectory for the root of the robot $R$, we are able to generate a discrete sequence of contact configurations  describing a feasible motion for $R$, between a start an goal configurations. The contact configurations are chosen to be statically balanced and task efficient, which means that they can generate a force  contributing to the motion of the root along the trajectory.

>>>>

\section{From a guide path to a discrete sequence of contact configurations ($\mathcal{P}_2$)}
\label{sec:contact}
As an input of this stage, we consider a truly feasible guide path $\mathbf{q}^0(t) : [0,1] \longrightarrow SE(3)$ for the root of the robot $R$. We now consider the second problem of computing a discrete sequence of contact configurations $\mathbf{Q}^{\overline{0}}$ for the limbs of the robot along $\mathbf{q}^0(t)$. From true feasibility we know that such a trajectory exists.

@@

In what follows a \textit{valid} contact configuration is understood as being collision free, respecting joint limits, and allowing to maintain the quasi static balance
of the robot. The first two tests are trivial; our criterion to assert efficiently the quasi static balance of the system
is described in Section TODO.

\subsection{Definition of a contact generator}
\label{sec:single_contact}
We consider a limb $R^k$  and a root configuration $\mathbf{q}^0$ truly feasible regarding $R^k$. FIGURE
From true feasibility we know that there exists an infinite list of mappings $P = [\pi_0,..,\pi_i, ...]$ that verify Eq.~\eqref{eq:pi} for $\mathbf{q}^0$.
The issue is then to evaluate locally at least one element of $P$ at $\mathbf{q}^0$.

While exhibiting analytically a $\pi_i$ does not seem tractable, we can iteratively evaluate $\pi_i(\mathbf{q}^0)$ as follows:
\begin{enumerate}
\item Generate randomly a valid limb configuration;
\item Project the end-effector onto the closest surface with inverse kinematics;
\item If a valid solution is found, stop. Otherwise repeat from step 1.
\end{enumerate}

We partially trade the completeness property of this approach for a more efficient solution previously introduced \citep{Tonneau2014}, that we describe
here for completeness.

We first define $C_{Contact}^{\epsilon} \supset C_{Contact}$ as the set of configurations almost in contact. This means that the minimal distance 
between an effector and a surface of the environment is lesser than a small value $\epsilon$.
We then apply the following steps:
\begin{enumerate}
\item Generate offline $N$ valid sample limb configurations $\mathbf{q}^k_i,  0 \leq i < N$;
\item Using the end-effector positions $\mathbf{p}(\mathbf{q}^k_i)$ as indices, store each sample in an octree stata structure;
\item Figure~\ref{fig:contact_gen}--2 and 3--: At runtime, when contact creation is required, retrieve from the octree the list of samples $S \subset C_{Contact}^{\epsilon}$ close to contact;
\item According to a user-defined heuristic $h$, sort the elements of $S$;
\item Figure~\ref{fig:contact_gen}--4 and 5--: Select the first configuration of $S$ (or fail if $S$ is empty). Project the configuration onto contact with inverse kinematics;
\item If a valid solution is found, stop. Otherwise remove the element from $S$ and go back step 5.
\end{enumerate}

This approach has two main advantages:
First, this allows us to select a large number of candidates in a single proximity request.
Having several candidates is interesting, because it allows to compare them regarding a user selected heuristic $h$, thus obtaining
a locally optimal candidate.
Furthermore, the fact that the candidates are already close to contact increases the odds that the inverse kinematics will converge to a valid solution,
and this in a small number of iterations.
Regarding completeness, it is immediate to verify that as $N$ and $\epsilon$ grow, the completeness of the method is restored.
$N$ is a parameter that allows us to specify the trade-off between completeness and efficiency.
The reader is invited to refer to our previous work for an extensive discussion on the optimal value of $N$ \citep{Tonneau2014}.

We thus have defined a real time, extremely efficient contact generator, that can be as complete as required by the context.
%We compute the image of a mapping $f$ along $\mathbf{q}^0(t)$ into a discrete sequence of contacts $ \in C_{contact}$. We first describe the discretization of $\mathbf{q}^0(t)$ and the algorithm used to create or break contacts along it. Then we describe how contacts are explicitely generated along the discretized trajectory.

\subsection{Extension of the guide trajectory}
Using the contact generator, we define an iterative algorithm to generate the contact
sequence along the guide path.
Although we address acyclic contact sequences, the algorithm is deterministic in the order in which 
the contacts are created, allowing it to break the combinatorial.
The complete algorithm can be found in Appendix~\ref{app:contact}.
In the remainder of the section we provide an intuition of it.

As an input, we consider the truly feasible guide path $\mathbf{q}^0(t)$, automatically discretized into a sequence of $j$ key placements:  
\begin{equation*}
	\mathbf{Q}^0 = [\mathbf{q}^0_{0}; \mathbf{q}^0_{i}; ..., \mathbf{q}^0_{j-1}]
\end{equation*} 
where $\mathbf{q}^0_{0}$ and $\mathbf{q}^0_{j-1}$ respectively correspond to the start and goal configurations. To ensure continuity in the contact transition phases, we rewrite $\pi$ under the following recursive form for any $0<i<j$:
\begin{equation*}
    \pi\colon\left\{
    \begin{aligned}		
        \mathbf{Q}^0 \in C_{reach} & \longrightarrow C_{contact} \\
        %\mathbf{q}^{0}_0 &  \longrightarrow  \mathbf{q}_{start} \\
        \mathbf{q}^{0}_i &  \longrightarrow  \mathbf{q}^{0}_i \oplus g(\mathbf{q}_{i - 1},\mathbf{q}^{0}_i) 
    \end{aligned}
    \right.
\end{equation*} 
We initialize the recurrence with $\pi(\mathbf{q}^0_{0}) = \mathbf{q}_0$ the initial configuration of the robot.

At each step, we consider a new root placement $\mathbf{q}^0_{i}$ and try to generate a balanced contact configuration that preserves as much as possible
the previous contacts.

%~ The function $g$ is defined independently by $g^k$ for each limb $R^k$. In defining $g^k$, two aspects must be considered. Is the limb $R^k$ in contact? And which criteria is it optimizing?

\subsubsection{Maintaining a contact in the sequence.}

Figure~\ref{fig:break_contact} illustrates the contact persistence method.
If possible, a limb in contact at time $i-1$ remains in contact at $i$. The contact is broken if an inverse kinematics solver fails to find a collision free limb configuration that satisfies joint limits. The solver is directy provided by the Hpp software.

If the solver fails, the contact is broken and a collision free configuration is assigned to the limb.
If more than one contacts are broken for a single step, intermediate steps are added to reposition, one at a time,
the contacts, before moving.

\begin{figure}[t]
\centering
  \begin{overpic}[width=0.9\linewidth]{figures/break_contact}
		\put (0,4) {1} 
		\put (25,4) {2} 
		\put (50,4) {3} 
		\put (76,4) {4} 
		%~ \put (68,58) {3.a)} 
		%~ \put (5,27) {3.b)} 
		%~ \put (37,27) {4.a)} 
		%~ \put (68,27) {4.b)} 
	\end{overpic}
\caption{Contacts are maintained if joint limits and collisions constraints are respected (2). They are broken otherwise(3,4).}
		   \label{fig:break_contact}
\end{figure}

%~ \begin{figure}[t]
%~ \centering
  %~ \begin{overpic}[width=0.6\linewidth]{figures/generate_contact}
		%~ \put (5,58) {1)} 
		%~ \put (37,58) {2)} 
		%~ \put (68,58) {3.a)} 
		%~ \put (5,27) {3.b)} 
		%~ \put (37,27) {4.a)} 
		%~ \put (68,27) {4.b)} 
	%~ \end{overpic}
%~ \caption{Contacts are generated when the configuration is not balanced.}
		   %~ \label{fig:generate_contact}
%~ \end{figure}

%\subsection{Generation of a contact configuration}  
\subsection{Creating contacts.}
Contacts are created using three rules:
\begin{enumerate}
\item Only one contact creation can happen between two consecutive states. This increases the odds that the interpolation between the two states will be feasible; 
\item A contact is validated if and only if the resulting configuration is statically balanced, no matter what user-defined heuristics are used; 
\item We use a FIFO approach:  we always try to first create a contact with the limb that has been free for the longest time. If the contact creation
was not successful for a limb, the limb is pushed on top of the queue, and will only be tried again after the others have been treated.
\end{enumerate}

These three rules provide a deterministic manner to generate contacts along the discretized guide path.
FIGURE. For a given truly feasible path, our current implementation does not guarantee that the planner will succeed in generating a balanced configuration, because
of the deterministic approach used to break the combinatorial.
However, in practice the planner is successful in the large majority of cases, as discussed in section~\ref{sec:stat1}, thanks to 
the heuristics we use for relevant contact selection.

### 2015-12-02 body/equilibrium.tex
finished contact generation now heuristics

\section{From a guide trajectory to a discrete sequence of contact configurations (Stage 2)}
\label{sec:efort}
As an input of this stage, we consider a truly feasible guide trajectory $\mathbf{q}^0(t) : [0,1] \longrightarrow SE(3)$ for the root of the robot $R$. We now consider the second problem of computing a trajectory $\mathbf{q}^{\overline{0}}(t)$ for the limbs of the robot. Since we assume true feasibility, we know that such a trajectory exists. Contrary to previous works \cite{DBLP:conf/iser/EscandeKMG08,Bouyarmane2009}, the goal here is not to find any such trajectory but rather to select one with particular properties. Specifically, we show here how to build a contact sequence with a small number of contact variations and  good-efficiency and naturalness of the postures.

More precisely, any mapping $\pi$ introduced in \eqref{eq:pi} can be used to expand $\mathbf{q}^0(t)$ into a whole-body trajectory. We propose here a particular construction of $\pi$ leading to interesting contact sequences.

%We compute the image of a mapping $f$ along $\mathbf{q}^0(t)$ into a discrete sequence of contacts $ \in C_{contact}$. We first describe the discretization of $\mathbf{q}^0(t)$ and the algorithm used to create or break contacts along it. Then we describe how contacts are explicitely generated along the discretized trajectory.

\subsection{Extension of the guide trajectory} 
The guide trajectory $\mathbf{q}^0(t)$ is first discretized into a sequence of $j$ key placements:  
\begin{equation*}
	\mathbf{Q}^0 = [\mathbf{q}^0_{0}; \mathbf{q}^0_{i}; ..., \mathbf{q}^0_{j-1}]
\end{equation*} 
where $\mathbf{q}^0_{0}$ and $\mathbf{q}^0_{j-1}$ respectively correspond to the start and goal configurations.  To ensure continuity in the contact transition phases, we rewrite $\pi$ under the following recursive form for any $0<i<j$:
\begin{equation*}
    \pi\colon\left\{
    \begin{aligned}		
        \mathbf{Q}^0 \in C_{reach} & \longrightarrow C_{contact} \\
        %\mathbf{q}^{0}_0 &  \longrightarrow  \mathbf{q}_{start} \\
        \mathbf{q}^{0}_i &  \longrightarrow  \mathbf{q}^{0}_i \oplus g(\mathbf{q}_{i - 1},\mathbf{q}^{0}_i) 
    \end{aligned}
    \right.
\end{equation*} 
We initialize the recurrence with $\pi(\mathbf{q}^0_{0}) = \mathbf{q}_0$ the initial configuration of the robot.
The function $g$ is defined independently by $g^k$ for each limb $R^k$. In defining $g^k$, two aspects must be considered. Is the limb $R^k$ in contact? And which criteria is it optimizing?

\medskip%\subsection{Maintaining and breaking contact in the sequence}
\noindent\textbf{Maintaining a contact:} If possible, a limb in contact at time $i-1$ remains in contact at $i$. The contact is broken if an inverse kinematics solver fails to find a collision free limb configuration which satisfies joint limits.
The solver is a direct implementation of the inverse kinematics solver with task priorities framework proposed in \cite{Baerlocher2004}, which
handles explicitely joint limits, and allows to integrate collision avoidance as a secondary criterion.

If the solver fails, the contact is broken and a collision free configuration is assigned to the limb.
%~ The configuration is selected to be the closest to the previous configuration (by minimizing $\| \mathbf{p}^k(\mathbf{q_i}) - \mathbf{p}^k(\mathbf{q_{i-1}}) \|$).

Once a first candidate configuration is taken for all limbs, the quasi-static balance is tested by whether the weight wrench is in the gravito-inertial cone (i.e. there exists valid contact forces that compensate for the weight of the robot), using the geometric approach described in \cite{qiu:dhm:2011}. If the balance is not obtained, new contacts are randomly generated using the following procedure.

%\begin{itemize} 
%\item If the robot is not statically balanced and $\mathbf{q}_{i-1}^k$ is not in contact, $g$ computes a configuration $\mathbf{q}_{i}^k$ so that a contact is taken with the environment (Fig.~\ref{fig:generate_contact}).
%\item Else, $g$ computes a configuration $\mathbf{q}_{i}^k$ minimizing $\| \mathbf{p}^k(\mathbf{q_i}) - \mathbf{p}^k(\mathbf{q_{i-1}}) \|$, subject to $R^k(\mathbf{q_i}) \cap O = \emptyset$ and $\mathbf{q_i}^k$ respects the  joint limits of $R^k$.
%
%If $\mathbf{q}_{i - 1}^k$ is in contact, this results in trying to maintain the contact for $\mathbf{q}_{i}^k$ (Figure~\ref{fig:break_contact}--1 and 2). The minimum is different than zero if $W^k(\mathbf{q}_i^0) \cap  \mathbf{p}^k(\mathbf{q_{i-1}}) = \emptyset$ (Figure~\ref{fig:break_contact}--3), or collision can not be  prevented (Figure~\ref{fig:break_contact}--4).
%\end{itemize}

\begin{figure}[t]
\centering
  \begin{overpic}[width=0.6\linewidth]{figures/break_contact}
		\put (3,15) {1)} 
		\put (27,15) {2)} 
		\put (50,15) {3)} 
		\put (75,15) {4)} 
		%~ \put (68,58) {3.a)} 
		%~ \put (5,27) {3.b)} 
		%~ \put (37,27) {4.a)} 
		%~ \put (68,27) {4.b)} 
	\end{overpic}
\caption{Contacts are maintained unless their position is too far, or the environment prevents it.}
		   \label{fig:break_contact}
\end{figure}

\begin{figure}[t]
\centering
  \begin{overpic}[width=0.6\linewidth]{figures/generate_contact}
		%~ \put (5,58) {1)} 
		%~ \put (37,58) {2)} 
		%~ \put (68,58) {3.a)} 
		%~ \put (5,27) {3.b)} 
		%~ \put (37,27) {4.a)} 
		%~ \put (68,27) {4.b)} 
	\end{overpic}
\caption{Contacts are generated when the configuration is not balanced.}
		   \label{fig:generate_contact}
\end{figure}

%\subsection{Generation of a contact configuration}  
\medskip\noindent\textbf{Creating a contact:} We consider a configuration where some limbs are in contact, some are free and quasi-static balance is not enforced. To enforce balance, we proceed in the following 
manner:
we randomly select a contact free limb; if there is no contact free limb, we select
the limb that made contact first.
Using the contact generator introduced in~\cite{Tonneau2014}, we project the configuration of this limb into a contact that enhances balance, if it exists (Figure~\ref{fig:generate_contact}); If balance is not achievable and a contact is possible, it is generated anyway;
If balance is not achieved, the next limb is selected and projected into a contact configuration, and so on.
This approach can lead to the repositioning of existing contacts, in which case intermediate states are inserted to reposition the contacts.
This current implementation does not guarantee that the planner will succeed in generating a balanced configuration, because
true feasibility is not fully guaranteed.
However in practice the planner is successful in the large majority of cases, as discussed in section~\ref{sec:stat1}.
%~ A new contact configuration is obtained by projecting a random configuration to the nearest contacts. The sampling is biased by selecting configurations so that the limbs previously in contact are already close to contact and at least one additional limb is close to the environment. The random configuration is then projected onto the contact by iteratively minimizing the distance between the end effectors and the environment, under the constraint of respecting the joint limits and avoiding the collision.

\subsection{Contribution to the global movement: the EFORT criteria}
\noindent\textbf{EFORT criterion:} If only relying on the random sampling to select new contacts, the planner produces inefficient postures. The resulting contact sequence is then poorly efficient and unnatural. Moreover, the limbs are not well configured and are not able to efficiently follow the general movement: contacts break frequently.

>>>>

\section{Heuristics for relevant contact selection}

\subsection{A heuristic for robust quasi static balance}
\noindent\textbf{EFORT criterion:} If only relying on the random sampling to select new contacts, the planner produces inefficient postures. The resulting contact sequence is then poorly efficient and unnatural. Moreover, the limbs are not well configured and are not able to efficiently follow the general movement: contacts break frequently.

### 2015-12-02 body/intro.tex
finished contact generation now heuristics

>>>>

\begin{figure*}
  \centering
  \begin{overpic}[width=0.8\linewidth]{figures/workflow}
    %~ \put (6.4,1.8) {\normalsize{Request}} 
    \put (30,10.8) {\large{\color{white}\Pa} }
    %~ \put (31.6,2) {\scriptsize{\color{white}RB-RRT}} 
    \put (66.4,10.8) {\large{\color{white}\Pb} }
    %~ \put (72.5,2) {\tiny{\color{white}Contact}} 
  \end{overpic}
  \vspace{-1em}
  \caption{
    Overview of our 2-stage framework. \Pa: Given a path request between the yellow and blue positions, a guide path is computed in the space of truly feasible guides $C_{reach}$. This is achieved by defining a geometric condition, the reachability condition (abstracted here with the transparent cylinders). \Pb: The trajectory is extended into a discrete sequence of contact configurations using an iterative algorithm.}
  \label{fig:framework}
\end{figure*}

### 2015-12-02 body/overview.tex
finished contact generation now heuristics

%
\begin{figure}
  \begin{overpic}[width=0.9\linewidth]{figures/contact_gen}
		\put (3,3) {1)} 
		\put (25,3) {2)} 
		\put (44,3) {3)} 
		\put (63,3) {4)} 
		\put (82,3) {5)} 
	\end{overpic}
  \caption{Generation of a contact configuration for the right arm of a humanoid robot. 1) Selection of reachable obstacles. 2) A request is performed on a database of configurations. 3) Configurations too far from contact are eliminated. 4) The best candidate according to EFORT is chosen. 5) The final contact is achieved using inverse kinematics.}
  \label{fig:contact_gen}
\end{figure}
\subsection{Computation of a guide path}
We first consider the problem of planning a relevant guide path. The objective is to compute a path of root placements which will allow contact creation. A desired requirement is to preserve the completeness of the planner: it should be able to explore any possible guide trajectory, but at the same time, any computed guide trajectory must be truly feasible, i.e. must lead to a valid sequence of contacts.

 An intuitive description of such placements is "close, but not too close": close, because a contact surface must be partially included in the Reachable workspace of the robot (represented for the right arm in Fig.~\ref{fig:contact_gen}--1); not too close, because the robot must avoid collision (which is represented by the hull including the torso in Fig.~\ref{fig:contact_gen}--1). We define formally $C_{reach}$, the set of interesting root placements, in which we compute a guide trajectory with a sampling based planner,  the Reachability-Based RRT --RB-RRT-- (Figure~\ref{fig:framework}--A). Planning in $C_{reach}$ boils down to planning in SE(3), which has an acceptable practical complexity.
%
Details are presented in Section~\ref{rbprm}.

\subsection{Generating a discrete sequence of contact configurations}

The second stage extends the guide path into a sequence of contact configurations (Fig.~\ref{fig:framework}--B). The true feasibility of the input path is guaranteed by the first step.
This provides in turn a reasonnable guarantee that a contact configuration can be obtained for any root placement along the path. With a fixed root position, the dimensionality of the 
contact generation problem is reduced to the number of degrees of freedom of the considered limb. In other words, we consider each limb as a manipulator attached to the root, and select the most relevant contact from a database of precomputed configurations, independent from the environment (Figure~\ref{fig:contact_gen}).

@@

\begin{figure}[][t]
  \centering
  \begin{overpic}[width=0.46\linewidth]{figures/character}
    \put (7,29) {\tiny{$\in \mathbf{q}^1$}}
    \put (7,25) {\tiny{effector}}
  \end{overpic}
  \caption{
    Left: Robot in a rest configuration. The right arm is denoted as the limb $R^1$. Each colored dot represents a degree of freedom around an axis. Right: Volumes of the robot. The red geometry denotes $W^0$ and must remain collision-free. The green spheres are the  $W^k$.}
  \label{fig:character}
\end{figure}

>>>>

Figure~\ref{fig:framework} illustrates our workflow.
\Pa and \Pb are addressed in a sequential fashion.
The guarantees provided by the truly feasible guide path allow
us to avoid the combinatorial by generating the contacts in a straightforward manner.

%
\begin{figure*}
  \centering
  \begin{overpic}[width=0.8\linewidth]{figures/contact_gen}
		\put (1,1) {1} 
		\put (22,1) {2} 
		\put (42,1) {3} 
		\put (62,1) {4} 
		\put (83,1) {5} 
	\end{overpic}
  \caption{Generation of a contact configuration for the right leg of HRP-2. 1: Selection of reachable obstacles. 2: Entries of the limb samples database (with $N = 4$). 3: With a proximity query on the octree database, configurations too far from obstacles are eliminated. 4: The best candidate according to a user-defined heuristic $h$ is chosen. 5: The final contact is achieved using inverse kinematics.}
  \label{fig:contact_gen}
\end{figure*}
\subsection{Computation of a guide path -- \Pa}
We first consider the problem of planning a relevant guide path. The objective is to compute a path of root placements which will allow contact creation. A desired requirement is to preserve the completeness of the planner: it should be able to explore any possible guide trajectory, but at the same time, any computed guide trajectory must be truly feasible, i.e. must lead to a valid sequence of contacts.

 An intuitive description of such placements is "close, but not too close": close, because a contact surface must be partially included in the Reachable workspace of the robot (represented for the right arm in Fig.~\ref{fig:contact_gen}--1); not too close, because the robot must avoid collision (which is represented by the hull including the torso in Fig.~\ref{fig:contact_gen}--1). We define formally $C_{reach}$, the set of interesting root placements, in which we compute a guide trajectory with a sampling based planner,  the Reachability-Based RRT --RB-RRT-- (Figure~\ref{fig:framework}--\Pa). Planning in $C_{reach}$ boils down to planning in SE(3), which has an acceptable practical complexity.
%
Details are presented in Section~\ref{rbprm}.

\subsection{Generating a discrete sequence of contact configurations -- \Pb}

The second stage extends the guide path into a sequence of contact configurations (Fig.~\ref{fig:framework}--\Pb). The true feasibility of the input path is guaranteed by the first step.
This provides in turn a reasonnable guarantee that a contact configuration can be obtained for any root placement along the path. With a fixed root position, the dimensionality of the 
contact generation problem is reduced to the number of degrees of freedom of the considered limb. In other words, we consider each limb as a manipulator attached to the root, and select the most relevant contact from a database of precomputed configurations, independent from the environment (Figure~\ref{fig:contact_gen}).

@@

\begin{figure}
  \centering
  \begin{overpic}[width=0.8\linewidth]{figures/character}
    \put (7,29) {\small{$\in \mathbf{q}^1$}}
    \put (7,25) {\small{effector}}
  \end{overpic}
  \caption{
    Left: Robot in a rest configuration. The right arm is denoted as the limb $R^1$. Each colored dot represents a degree of freedom around an axis. Right: Volumes of the robot. The red geometry denotes $W^0$ and must remain collision-free. The green spheres are the reachable workspace of each limb, the  $W^k$.}
  \label{fig:character}
\end{figure}
